# Reproducible Research: Peer Assessment 1

```{r}
library(data.table)
```
## Loading and preprocessing the data
```{r}
data<-fread("activity.csv")
```



## What is mean total number of steps taken per day?
```{r}
hist(data[,list(sum=sum(steps,na.rm=TRUE)),by=date]$sum,xlab="Number of steps",main="Number of steps taken each day")
```

Mean number of steps taken each day:
```{r}
mean(data[,list(sum=sum(steps,na.rm=TRUE)),by=date]$sum,na.rm=TRUE)
```

Median number of steps taken each day:
```{r}
median(data[,list(sum=sum(steps,na.rm=TRUE)),by=date]$sum,na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r}
averageInterval<-data[,list(sum=sum(steps,na.rm=TRUE)/61),by=interval]

plot(averageInterval$sum,type="l",xlab="5 min interval",ylab="Number of steps (mean of all days",main="Average number of steps in 5 min interval")
```

Interval that on average has the highest number of steps
```{r}
averageInterval[which.max(averageInterval$sum),interval]
```

## Imputing missing values
```{r}
missingValues<-!complete.cases(data$step)
```
Number of missing values in the dataset:
```{r}
sum(missingValues)
```
Impute missing values by taking the median of step numbers for each interval:

```{r}
imputedData<-data
imputedData$steps<-as.double(imputedData$steps)

for (i in 1:length(missingValues))
        {
                if(missingValues[i]){
                        imputedData[i,steps:=median(data[interval==data[i,interval],steps],na.rm=TRUE)]
                }
        }

hist(imputedData[,list(sum=sum(steps,na.rm=TRUE)),by=date]$sum,xlab="Number of steps",main="Number of steps taken each day")
```

Mean and median total number of steps after imputing missing values.
```{r}
mean(imputedData[,list(sum=sum(steps)),by=date]$sum)
median(imputedData[,list(sum=sum(steps)),by=date]$sum)
```
The mean is slightly higher than with the non-imputed data, the median is the same because we replaced the missing values by the median for each interval.
## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
```
Add a day column to the data frame, TRUE means weekend, FALSE is weekday:
```{r}
imputedData[,days:=weekdays(as.Date(data$date)) %in% c('Saturday','Sunday')]
```

Tidy up the new data and make the plot:
```{r}
averageImputed<-imputedData[,list(mean=mean(steps)),by="days,interval"]
averageImputed$days<-as.factor(averageImputed$days)
levels(averageImputed$days) <- c("Weekday","Weekend")
levels(averageImputed$days)

xyplot(mean~interval|days,averageImputed,  t="l",ylab="Number of steps",xlab="Interval")
```

